Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
Minor Edit

---
title: "Will Kamala Harris will win the 2024 Presidential election"
subtitle: "Analysing the polls using baysian modeling"
author: 
  - Yan Mezhiborsky
  - Cristina Burca
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
date: today
date-format: long
abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina. USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world. Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris. Provided this we predict that Kamal Harris will win the election. By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction. "
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

@@

In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER]. As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states. While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.

@@

# Plot the second subset
ggplot(subset2, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date") +
  theme_minimal() +
      scale_x_date(
    date_breaks = "1 month",
    date_minor_breaks = "1 week",
    date_labels = "%b %d" 
  ) +
   theme(legend.position = "none")
```
@fig-supportAM further examines Harris's support trends in Arizona and Michigan. Both states exhibit an upward trend in support starting in early October, which may suggest increasing favorability or strategic campaign efforts in these regions. This upward trajectory could indicate an opportunity for Harris to solidify her support in these competitive areas.

@@

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Pollster}_i + \beta_3 \times \mbox{Days since first poll}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```

The outcome variable in this model is the proportion of respondents who support Harris in each poll, represented as $y_i$. This binomial outcome (successes out of the total sample size) is modeled as a function of several predictor variables that account for differences by state, pollster, and time.

@@

```
In @fig-modelNPW, we see states of Nevada, Pennsylvania and Wisconsin have a singular declining trend.

@@

kable(predicted_support_table, col.names = c("State", "Avg. Predicted Support for Harris"), digits = 2)
```
```{r}

@@

Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.

@@

Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent. 
Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data. This means we will be spending this money to create the survey, and have a infrastructure in place to hold the large amount of data as well as cover any security fees to keep this data private and safe. Then we allocate $50000 to advertise the survey. We will need a large sample so we will be spending most of our budget for this. We will have a different urls for each advertisement we have so that we can incorporate this in our data to see which demographic is accessing the survey through which platform. The platforms to advertise this is through spotify, facebook, instagram, and some news networks. If possible we will try to use this money to get it endorsed by branches of government to show our reliability. Reliability is also what citizens respond to so this will get a lot of respondents. We will spend \$25,000 for modeling our data as they tend to be expensive with data this size. And the rest of the \$15,000 will be leftover cost for anything that we don't foresee. If there is any leftover we can add a survey participation price such as having a free 3 month trial for the platforms we mentioned above (that is if they have a subscription based membership).

>>>>

---
title: "Will Kamala Harris will win the 2024 Presidential election"
subtitle: "Analysing the polls using baysian modeling"
author: 
  - Yan Mezhiborsky
  - Cristina Burca
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
date: today
date-format: long
abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina. USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of the USA will help understand the future economic condiotion of the world. Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris. Provided this we predict that Kamal Harris will win the election. By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction. "
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

@@

In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin \[CITE NYT PAPER\]. As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states. While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.

@@

# Plot the second subset
ggplot(subset2, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date") +
  theme_minimal() +
      scale_x_date(
    date_breaks = "1 month",
    date_minor_breaks = "1 week",
    date_labels = "%b %d" 
  ) +
   theme(legend.position = "none")
```

@fig-supportAM further examines Harris's support trends in Arizona and Michigan. Both states exhibit an upward trend in support starting in early October, which may suggest increasing favorability or strategic campaign efforts in these regions. This upward trajectory could indicate an opportunity for Harris to solidify her support in these competitive areas.

@@

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Pollster}_i + \beta_3 \times \mbox{Days since first poll}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
The outcome variable in this model is the proportion of respondents who support Harris in each poll, represented as $y_i$. This binomial outcome (successes out of the total sample size) is modeled as a function of several predictor variables that account for differences by state, pollster, and time.

@@

In @fig-modelNPW, we see states of Nevada, Pennsylvania and Wisconsin have a singular declining trend.

@@

kable(predicted_support_table, col.names = c("State", "Avg. Predicted Support for Harris"), digits = 2)
```

```{r}

@@

Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.\@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.

@@

Here we will be talking about how we would conduct a \$100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent. Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data. This means we will be spending this money to create the survey, and have a infrastructure in place to hold the large amount of data as well as cover any security fees to keep this data private and safe. Then we allocate \$50000 to advertise the survey. We will need a large sample so we will be spending most of our budget for this. We will have a different urls for each advertisement we have so that we can incorporate this in our data to see which demographic is accessing the survey through which platform. The platforms to advertise this is through spotify, facebook, instagram, and some news networks. If possible we will try to use this money to get it endorsed by branches of government to show our reliability. Reliability is also what citizens respond to so this will get a lot of respondents. We will spend \$25,000 for modeling our data as they tend to be expensive with data this size. And the rest of the \$15,000 will be leftover cost for anything that we don't foresee. If there is any leftover we can add a survey participation price such as having a free 3 month trial for the platforms we mentioned above (that is if they have a subscription based membership).
